A Detection of the Environmental Dependence of the Stellar Profiles and haloes of Massive Central Galaxies

### 2017-07-13 draft/paper2.tex
Edit summary

        \item a 10 kpc radii (hereafter noted \minn{}) as proxy of the ``in-situ'' 
            stellar component. 
            This is motivated by recent observations and simulations 
            (e.g.~\citealt{vanDokkum2010}, \citealt{RodriguezGomez2016}).

@@

    We match the \rbcg{} and \nbcg{} samples using both \mtot{} and \minn{} in this
    work. 
    As explained in Paper~1, we use \mtot{} as proxy of the ``total'' \mstar{}
    of massive galaxies while treat \minn{} as a rough indicator of the mass of the 
    \textit{in-situ} component.

@@

    Generally speaking, such subtle environmental dependence of \mden{} profiles in 
    the \mtot{}-matched samples suggests that environment does not play a major role 
    in shaping massive galaxies \textbf{at fixed ``total'' stellar mass}.
    Therefore it is interesting to ask, for massive galaxies that \textbf{contain 
    similar \mstar{} in the \textit{in-situ} component}, how does environment affect 
    their subsequent mass assembly?
    Assuming \minn{} is reasonable tracer of the ``in-situ'' mass, we compare the 
    \mden{} profiles of \minn{}-matched samples (right panel of Fig~\ref{fig:prof_1}).  
    Although the median \mden{} profiles are very similar at $R \leq 10$ kpc, the
    massive central galaxies in more massive haloes host much more prominent outer 
    stellar halo.

@@

    We qualitatively compare our results with \citealt{HCompany13} (HC13; 
    green solid line) who studied $z\leq 0.09$ central ETGs in 
    $12.5 \le$ \logmh{} $< 15$ haloes. 
    HC13 uses the group catalog by \citet{Yang2007} to estimate \mhalo{}. 
    They estimate the 2-D $R_{\mathrm{e}}$ using single-\ser{} model fitting to SDSS 
    images, and derived \mstar{} based on SED fitting using the BC03 (\citealt{BC03}) 
    synthetic population model. 
    The \mstar{} estimates are empirically converted from a Kroupa IMF to a Chabrier 
    IMF by applying a constant -0.05 shift (see \citealt{Bernardi2016}).
    We add $+0.1$dex to the HSC13 \mstar{} to account for the systematic 
    difference between their BC03 and our FSPS model (see Appendix~\ref{app:sed}). 
    The distributions of both \rbcg{} and \nbcg{} galaxies on the 
    \mtot{}--$R_{\mathrm{50}}$ plane follow the HC13 relation reasonably well 
    (with slightly shallower slopes), even at \logmtot{}$< 11.6$ where our samples 
    start to become incomplete. 
    In contrast to the present study, HC13 finds no environment dependence in the 
    mass--size relation for central galaxies from $12.5\le$ \logmh{} $<15.0$ haloes.
    The \mstar{} and $R_{\mathrm{e}}$ in HC13 are based on much shallower SDSS images
    that can not constrain the environment-sensitive outer stellar halos well. 
    Detailed analysis on this topic is beyond the scope of this work, but we briefly 
    demonstrate this using the \rbcg{} and \nbcg{} galaxies that overlap with the GAMA
    survey (please see Appendix~\ref{app:gama} for details). 
    We matched the two samples using the \mstar{} measured using single-\ser{} 2-D
    model by the GAMA team (\citealt{Kelvin2012}). 
    At fixed \mgama{}, the \rbcg{} galaxies display much more significant stellar 
    halo despite they have very similar inner \mden{} profiles.  
    This shows again that the deeper HSC images has unique advantage in studying the
    connection between the assembly of massive galaxies and their dark matter haloes.

@@

        \item Once matched at fixed \mtot{}, the \rbcg{} and \nbcg{} galaxies 
            show similar average ellipticity profiles. 
            There is hint that suggests the central galaxies from more massive 
            haloes become slightly more elongated at $R>50$ kpc than the ones 
            in less massive halos with similar \mtot{}. 
            However, the uncertainties in the current profiles make it not 
            conclusive.

        \item For the \minn{}-matched samples, the \rbcg{} galaxies show higher
            average ellipticity than the \nbcg{} ones at $R > 10$ kpc. 
            The average ellipticities in the inner region are very similar. 
            This further suggests that the more elongated outer stellar haloes 
            of massive halos relate to the accreted stars.

@@

    Our results are consistent with the picture that massive central galaxies form via 
    a ``two-phase'' scenario. 
    According to this theory, the cores of massive galaxies were formed at $z{\sim} 2$ 
    during an intense period of in-situ star formation. 
    The outskirts of massive galaxies are then built up via a more gradual second 
    phase of evolution (the ``ex-situ'' phase) which is dominated by mass growth via 
    the accretion of satellite systems. 
    Non-dissipative minor mergers mostly deposit stars on the outskirts of the 
    centrals and do not have large impact on the central \mden{} profile 
    (e.g. \citealt{Oogi2013, Bedorf2013}). 
    Hence, according to this picture, the outer stellar envelopes of massive centrals 
    should correlate with the masses of their host haloes, in qualitative agreement with
    our findings.

@@

    We now investigate to what degree our assumption that the inner 10 kpc \mstar{}
    corresponds to an in-situ mass.    
    Fig \ref{fig:discussion_1} compares the median \mden{} profiles of the 
    galaxies in two \mtot{} bins 
    ($11.4\leq$\logmtot{}$<11.6$ and $11.6\leq$\logmtot{}$<11.8$) with 
    (1) the median \mden{} profiles of massive ETGs at $1.0 < z < 1.5$ in
    \citealt{Patel2013} since they are considered as the progenitors of 
    ${\sim} 10^{11.5} M_{\odot}$ ETGs at $z=0$.    
    (2) the inner components of $z{\sim} 0$ ellipticals from 2-D decomposition 
    (\citealt{Huang2013a}).
    As proxy of the in-situ components, \citet{Huang2013b} showed they are 
    structurally similar to the compact ``red nuggets'' at high-$z$. 
    (3) the ``in-situ'' components of simulated central galaxies in massive haloes 
    from \citet{Cooper13} (the inner ${\sim} 5$ kpc is quite uncertain due to the 
    resolution).  
    The comparison confirms that \minn{} should be reasonable proxy of the 
    ``in-situ'' mass\footnote{We convert these \mden{} profiles to the same 
    Chabrier IMF; but there are still differences in median \mstar{} and 
    details in the \m2l{} estimates}.  
    More importantly, it shows that the \mden{} profile at 
    $\mathrm{R} > 15$-20 kpc is likely dominated by stars that are not related 
    to the ``in-situ'' processes.

@@

%% ------------------------------------------------------------------------------------ %%
   \song{This part is directly copied from Jenny's txt file, not the right place.}    
   \begin{itemize}

         \item At fixed \mstar{} evaluate within 100 kpc, we find that massive central 
            galaxies display \mden{} profiles that are not self-similar and which 
            present subtle but systematic differences that correlate with \mhalo{}.
            On average, massive central galaxies in more massive haloes have slightly 
            shallower inner \mden{} profiles and more prominent outer envelope.

        \item We also study how the \mden{} profiles vary with \mhalo{} at fixed 
            ``in-situ'' mass where it is approximated by the \mstar{} within 10 kpc. 
            At fixed \minn{} we find that central galaxies in more massive haloes 
            possess much more extended stellar envelope in the outskirts than the ones 
            in smaller haloes. 
            At the same time the ellipticity, and optical color profiles inside 10 kpc 
            show little to no dependence on environment.

        \item The \mstar{}-size relation also exhibits a clear environmental 
            dependence and central galaxies in more massive haloes have larger 
            $R_{\mathrm{50}}$ at fixed \mtot{}. 
            We suggest that the relation between \mtot{} and \minn{}\footnote{or any 
            better proxy of total \mstar{} and the mass formed in the ``in-situ'' phase.} 
            can be a useful tool in diagnosing the assembly history of these massive 
            galaxies and the connection with the assembly of their dark matter haloes.

@@

    \song{From old draft, need some works}

    In this work, we study how environment (halo mass) affects the stellar mass density 
    profiles of massive central galaxies using deep images from the Subaru HSC 
    survey. 
    With the help of this high-quality and wide area data set, we directly map the 
    stellar mass distributions of ${\sim}7000$ massive central galaxies at 
    $0.3 < z < 0.5$ out to $>100$ kpc without resorting to stacking techniques. 
    We group massive central galaxies into two categories based on their host halo 
    mass (\mhalo{}$\simgt 10^{14.2} M_{\odot}$ and \mhalo{}$\simlt 10^{14} M_{\odot}$). 
    Our main results are:

    \begin{enumerate}
        \item We find that the ``total'' \mstar{} of these massive galaxies can be 
            significantly underestimated with shallow imaging data such as SDSS and/or 
            imperfect model assumption (e.g. the \texttt{cModel} or single-\ser{}) 
            are used. 
            In contrast to previous work, our results do not depend on stacking or any 
            parametric models. 
            Moreover, the level of such underestimation could also depend on the 
            stellar mass and on halo mass. 
            This should be carefully taken into account when discussing topics such 
            as the evolution of the galaxy SMF.

        \item We show that the \mden{} profiles of massive galaxies are generally quite 
            smooth out to 10 kpc but that there is a large but continuous scatter in the 
            amount of light in their outskirts. 
            {\bf KEEP? We argue that ``cD'' galaxies simply represent the tail end of a 
            distribution but that they are not a distinct class of massive galaxies.}

        \item {\bf How do profiles depend on Mstar?}

@@

    These results highlight the advantages of wide area, deep, and high-quality imaging 
    for studying the evolution of massive galaxies. 
    Upon finishing this work, the HSC survey has already doubled its sky coverage to 
    ${\sim} 200$ deg$^2$, and provides a much larger sample of massive central galaxies. 
    In the near future, we will extend this work to lower \mtot{} by using photometric 
    redshifts, and we will also apply 2-D photometry method (e.g.\ \citealt{Huang2013a}) 
    to take advantages of the multi-wavelength nature of the HSC survey 
    (e.g. \citealt{Huang2016}). 
    Our current work can also be combined with weak lensing measurements of the dark 
    matter haloes of massive galaxies and physical insights into the assembly histories 
    of these galaxies can be gained by comparing with cosmological hydro-simulations 
    such as Illustris (\citealt{Vogelsberger2014}, \citealt{Genel2014}), 
    EAGLE (\citealt{Schaye2015}, \citealt{Crain2015}), or \textit{Horizon-AGN} 
    (\citealt{Dubois2014}).

>>>>

        \item a 10 kpc radii (hereafter noted \minn{}) as proxy of the ``in situ'' 
            stellar component. 
            This is motivated by recent observations and simulations 
            (e.g.~\citealt{vanDokkum2010}, \citealt{RodriguezGomez2016}).

@@

    We match the \rbcg{} and \nbcg{} samples using both \mtot{} and \minn{} in this
    work. 
    As explained in Paper~1, we use \mtot{} as proxy of the ``total'' \mstar{}
    of massive galaxies while treat \minn{} as a rough indicator of the mass of the 
    \textit{in situ} component.

@@

    Generally speaking, such subtle environmental dependence of \mden{} profiles in 
    the \mtot{}-matched samples suggests that environment does not play a major role 
    in shaping massive galaxies \textbf{at fixed ``total'' stellar mass}.
    Therefore it is interesting to ask, for massive galaxies that \textbf{contain 
    similar \mstar{} in the \textit{in situ} component}, how does environment affect 
    their subsequent mass assembly?
    Assuming \minn{} is reasonable tracer of the ``in situ'' mass, we compare the 
    \mden{} profiles of \minn{}-matched samples (right panel of Fig~\ref{fig:prof_1}).  
    Although the median \mden{} profiles are very similar at $R \leq 10$ kpc, the
    massive central galaxies in more massive haloes host much more prominent outer 
    stellar halo.

@@

    We qualitatively compare our results with \citealt{HCompany13} (HC13; 
    green solid line) who studied $z\leq 0.09$ central ETGs in 
    $12.5 \le$ \logmh{} $< 15$ haloes. 
    HC13 uses the group catalog by \citet{Yang2007} to estimate \mhalo{}. 
    They estimate the 2-D $R_{\mathrm{e}}$ using single-\ser{} model fitting to SDSS 
    images, and derived \mstar{} based on SED fitting using the BC03 (\citealt{BC03}) 
    synthetic population model. 
    The \mstar{} estimates are empirically converted from a Kroupa IMF to a Chabrier 
    IMF by applying a constant -0.05 shift (see \citealt{Bernardi2016}).
    We add $+0.1$dex to the HSC13 \mstar{} to account for the systematic 
    difference between their BC03 and our FSPS model (see Appendix~\ref{app:sed}). 
    The distributions of both \rbcg{} and \nbcg{} galaxies on the 
    \mtot{}--$R_{\mathrm{50}}$ plane follow the HC13 relation reasonably well 
    (with slightly shallower slopes), even at \logmtot{}$< 11.6$ where our samples 
    start to become incomplete. 
    In contrast to the present study, HC13 finds no environment dependence in the 
    mass--size relation for central galaxies from $12.5\le$ \logmh{} $<15.0$ haloes.
    The \mstar{} and $R_{\mathrm{e}}$ in HC13 are based on much shallower SDSS images
    that can not constrain the environment-sensitive outer stellar haloes well. 
    Detailed analysis on this topic is beyond the scope of this work, but we briefly 
    demonstrate this using the \rbcg{} and \nbcg{} galaxies that overlap with the GAMA
    survey (please see Appendix~\ref{app:gama} for details). 
    We matched the two samples using the \mstar{} measured using single-\ser{} 2-D
    model by the GAMA team (\citealt{Kelvin2012}). 
    At fixed \mgama{}, the \rbcg{} galaxies display much more significant stellar 
    halo despite they have very similar inner \mden{} profiles.  
    This shows again that the deeper HSC images has unique advantage in studying the
    connection between the assembly of massive galaxies and their dark matter haloes.

@@

        \item Once matched at fixed \mtot{}, the \rbcg{} and \nbcg{} galaxies 
            show similar average ellipticity profiles. 
            There is hint that suggests the central galaxies from more massive 
            haloes become slightly more elongated at $R>50$ kpc than the ones 
            in less massive haloes with similar \mtot{}. 
            However, the uncertainties in the current profiles make it not 
            conclusive.

        \item For the \minn{}-matched samples, the \rbcg{} galaxies show higher
            average ellipticity than the \nbcg{} ones at $R > 10$ kpc. 
            The average ellipticities in the inner region are very similar. 
            This further suggests that the more elongated outer stellar haloes 
            of massive haloes relate to the accreted stars.

@@

    Our results are consistent with the picture that massive central galaxies form via 
    a ``two-phase'' scenario. 
    According to this theory, the cores of massive galaxies were formed at $z{\sim} 2$ 
    during an intense period of in situ star formation. 
    The outskirts of massive galaxies are then built up via a more gradual second 
    phase of evolution (the ``ex-situ'' phase) which is dominated by mass growth via 
    the accretion of satellite systems. 
    Non-dissipative minor mergers mostly deposit stars on the outskirts of the 
    centrals and do not have large impact on the central \mden{} profile 
    (e.g. \citealt{Oogi2013, Bedorf2013}). 
    Hence, according to this picture, the outer stellar envelopes of massive centrals 
    should correlate with the masses of their host haloes, in qualitative agreement with
    our findings.

@@

    We now investigate to what degree our assumption that the inner 10 kpc \mstar{}
    corresponds to an in situ mass.    
    Fig \ref{fig:discussion_1} compares the median \mden{} profiles of the 
    galaxies in two \mtot{} bins 
    ($11.4\leq$\logmtot{}$<11.6$ and $11.6\leq$\logmtot{}$<11.8$) with 
    (1) the median \mden{} profiles of massive ETGs at $1.0 < z < 1.5$ in
    \citealt{Patel2013} since they are considered as the progenitors of 
    ${\sim} 10^{11.5} M_{\odot}$ ETGs at $z=0$.    
    (2) the inner components of $z{\sim} 0$ ellipticals from 2-D decomposition 
    (\citealt{Huang2013a}).
    As proxy of the in situ components, \citet{Huang2013b} showed they are 
    structurally similar to the compact ``red nuggets'' at high-$z$. 
    (3) the ``in situ'' components of simulated central galaxies in massive haloes 
    from \citet{Cooper13} (the inner ${\sim} 5$ kpc is quite uncertain due to the 
    resolution).  
    The comparison confirms that \minn{} should be reasonable proxy of the 
    ``in situ'' mass\footnote{We convert these \mden{} profiles to the same 
    Chabrier IMF; but there are still differences in median \mstar{} and 
    details in the \m2l{} estimates}.  
    More importantly, it shows that the \mden{} profile at 
    $\mathrm{R} > 15$-20 kpc is likely dominated by stars that are not related 
    to the ``in situ'' processes.

@@

@@

    Using the high-quality images from the first $\sim100$ deg$^2$ data of the 
    Hyper Suprime-Cam Subaru Strategic Program, we group a large sample of massive 
    central galaxies at $0.3 < z < 0.5$ into two categories based on their dark matter
    halo mass (\mhalo{}$\simgt 10^{14.2} M_{\odot}$ and 
    \mhalo{}$\simlt 10^{14} M_{\odot}$), and study how their stellar mass distributions 
    out to 100 kpc depend on their environments (or \mhalo{}). 
    The main results are:

        \item At fixed \mtot{}, we find that massive central galaxies display \mden{} 
            profiles that have large intrinsic scatter in the outer stellar haloes and 
            present subtle but systematic trend with \mhalo{}.
            On average, massive central galaxies in more massive haloes have slightly 
            shallower inner \mden{} profiles and more prominent outer haloes. 
            This suggests that the environment affects the assembly of massive central
            galaxies at fixed ``total'' stellar mass.

        \item At fixed \minn{}, massive central galaxies in more massive haloes 
            possess much more extended stellar haloes than the ones in smaller haloes, 
            despite their median \mden{} profiles at $R < 15$ kpc are very similar. 
            Assuming that \minn{} traces the amount of ``in situ'' stars in massive
            galaxies, this results is consistent with the picture that more massive 
            haloes favor the growth of extended stellar haloes in their central 
            galaxies through more frequent mergers.

        \item The \mtot{}-$R_{\mathrm{50}}$ relation for massive central galaxies 
            also exhibits environmental dependence. 
            Central galaxies in more massive haloes have slightly larger
            $R_{\mathrm{50}}$ than the ones from less massive haloes at fixed \mtot{}.

        \item We also find that massive central galaxies in different environments
            follow different relations between their \mtot{} and \minn{}. 
            Assuming that \mtot{} is a reasonable proxy of ``total'' stellar mass and 
            \minn{} reflects the amount of \textit{in situ} stars, we propose that 
            this relation can help diagnose the assembly of massive central galaxies 
            without using ``size'' with ambiguous definition.

        \item At fixed \mtot{}, the ellipticity profiles of massive central galaxies 
            do not display strong environment dependence at $R < 60$ kpc. 
            At fixed \minn{}, the more extended stellar haloes presented in the \rbcg{}
            galaxies also show higher ellipticity than the \nbcg{} galaxies.

        \item We do not find evidence that suggests the rest-frame color profiles 
            within 60 kpc depend on environment for neither \mtot{}- nor 
            \minn{}--matched samples.

@@

    These results highlight the importance of deep, high-quality images for studying 
    the assembly of massive haloes and their central galaxies. 
    We are currently working on combining the current results with weak lensing 
    measurements of dark matter haloes around these massive galaxies to gain more physical 
    insights. 
    We will also compare our results with cosmological hydro-simulations 
    such as 
    \textit{Illustris-TNG} (\citealt{Pillepich2017a, Pillepich2017b, Weinberger2017}, 
    \textit{Horizon-AGN} (\citealt{Dubois2014}), or 
    \textit{MassiveBlackII} (\citealt{Khandai2015}).
